PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-02-28 txt/manuscript.tex
Update CLI description (it's not yet finished)

We can now examine the results. We expect that for noisy simulations, the result will be further
away from the target value of \texttt{0.8535533905932737}. For example, in one of our runs we
obtained \texttt{prob\_succ\_noiseless = 0.8524401115559386} and \texttt{prob\_succ\_noisy =
0.5017958400693446}. 
\noindent
This concludes our example. In the next section, we will show how to use PyQBench's CLI.

\subsection{Using \texttt{qbench} CLI} 
Using PyQBench as a library allows one to conduct arbitrary two-qubit von Neumann measurement experiment. However, as discussed in the previous guide, it requires some amount of boilerplate code.
For a Fourier parametrized family of measurements, PyQBench offers a simplified way of performing the experiment using a Command Line Interface (CLI).
The workflow is describe as a list of steps:

@@

\item Preparing configuration files describing backend and the experiment scenario.
\item Submitting/running experiments. Depending on the experiment scenario, execution can be synchronous, or asynchronous.
\item (optional) Checking status of the submitted jobs if the execution is asynchronous.
\item For asynchronous jobs, 
resolve jobs into actual measurements.
\item Computing discrimination probabilities and outputting the resulting table.
\end{enumerate} 
\subsubsection{Preparing configuration files}
To prepare configuration files we need to define the experiment and the   backend 
which we want to benchmark. Experiments and backends are defined in YAML files. An example experiment definition looks as follows.
\begin{lstlisting}[language=Python, caption=Defining the experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}

The components of the experiment is as follows:
\begin{itemize}
	\item The \textit{type} describes type of the experiment. Currently, the only option of  \textit{type} is \textit{discrimination-fourier}. 
	\begin{lstlisting}[language=Python]
type: discrimination-fourier
	\end{lstlisting}
	\item The \textit{qubits} list describes pair of qubits on which the experiment should be run. In this example, the benchmark will run on two pairs of qubits. First pair is 0 and 1, second pair is 1 and 2.
\begin{lstlisting}[language=Python]
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
	  ancilla: 2
	\end{lstlisting}
\item The \textit{angles} key describe range of angles for Fourier parameterized family. The range is always uniform, starts at the \textit{start}, ends at \textit{stop} and contains \textit{num\_steps} points, including both start and stop. The start and stop can be arithmetic expressions using \textit{pi} literal. For instance range defined as:  
\begin{lstlisting}[language=Python]
angles:
	start: 0
	stop: 2 *pi
	num_steps: 3
\end{lstlisting}
 would contain three angles $0, \pi$  and $2\pi$.
 \item 
 The \textit{gateset} key describes the gate decomposition of the experiment. The possible options are \textit{[ibmq, lucy, rigetti, generic]}. If the user choose the device-dependent key, that is one of the options \textit{ibmq, lucy, rigetti}, then the experiment will be implemented by using predefined sequence of native gates. If the user choose the \textit{generic} key, then the compiler  will automatically decompose given circuit in the way that is suitable for the device in use.
 \begin{lstlisting}[language=Python]
gateset: ibmq
 \end{lstlisting}
 \item The \textit{method} can take one of two values: \textit{direct\_sum} or \textit{postselection}, and as the name suggests, defines which method is used to implement the experiment.
 \begin{lstlisting}[language=Python]
method: direct_sum
 \end{lstlisting}
 \item The \textit{num\_shots} defines how many shots are performed in the experiment for particular angle, qubit pair and circuit. For the purpose of calculating possible costs, this is how you compute total number of shots and executed circuits. Let $N$ be the number of qubit pairs in \textit{qubits} key and let $M$ be the number of steps defined in the \textit{angles} key. Then number of executed circuits is $2MN$ for the \textit{direct\_sum} method and $4MN$ for the \textit{postselection} method. Each of those circuits is executed \textit{num\_shots} times.
 \begin{lstlisting}[language=Python]
num_shots: 100
 \end{lstlisting}
\end{itemize}

Observe that the experiment file does not specify what backend to use. This way, the same experiment file can be used on multiple backends.
 Different backends may require different information to be used. For instance, IBMQ backends might require hub, group and project, whereas backends from the \textit{Aer} package do not. Hence, the exact format of the YAML file depends on what backend one wants to use. Below we present 
 YAML file defining the IMBQ backend named Quito. 
\begin{lstlisting}[language=Python, caption=IBMQ backend]
name: ibmq_quito
asynchronous: false
provider:
	hub: ibm-q
	group: open
	project: main
\end{lstlisting}

IBMQ backends typically require and access token to IBM Quantum Experience. It would be unsafe to store them in plain text, and therefore the token is configured separately. Before running the experiment, you should place your token in the \textit{IBMQ\_TOKEN} environmental variable.

\subsubsection{Notes on using the asynchronous flag}
In this example, the only key that requires the explanation is \textit{asynchronous} which determines whether experiments will be run asynchronously  (\textit{true}) or not (\textit{false}).
% We recommend setting it to true, unless your experiment is really small (several circuits total). However, some backends may not support asynchronous execution. This might be especially the case with local simulators.

For asynchronous execution to work, the following conditions have to be met:
\begin{itemize}
\item Jobs returned by the backend have unique \textit{job\_id}, 
\item Job object is retrievable from the backend using \textit{backend.retrieve\_job} method, even from another process (e.g. if the original process running the experiment has finished).
\end{itemize}
Since PyQBench cannot determine if the job retrieval works for given backend, the user is responsible for checking it yourself.

\todo[inline]{co z tymi sekcjami?}
\subsubsection{Backends obtainable from simple providers???}
\subsubsection{Backends constructed from a simple function call???}
\todo[inline]{co z tymi sekcjami?}

>>>>

We can now examine the results. As an example, in one of our runs, we obtained
\texttt{prob\_succ\_noiseless = 0.8524401115559386} and \texttt{prob\_succ\_noisy =
0.5017958400693446}. As expected, for noisy simulations, the result lies further
away from the target value of \texttt{0.8535533905932737}.

This concludes our example. In the next section, we will show how to use PyQBench's CLI.

\subsection{Using \texttt{qbench} CLI}
Using PyQBench as a library allows one to conduct a two--qubits benchmark with arbitrary von Neumann
measurement. However, as discussed in the previous guide, it requires writing some amount of code.
For a Fourier parametrized family of measurements, PyQBench offers a simplified way of conducting
benchmarks using a Command Line Interface (CLI). The workflow with PyQBench's CLI can be summarized
as the following list of steps::

@@

\item Preparing configuration files describing the backend and the experiment scenario.
\item Submitting/running experiments. Depending on the experiment scenario, execution can be synchronous, or asynchronous.
\item (optional) Checking the status of the submitted jobs if the execution is asynchronous.
\item Resolving asynchronous jobs into the actual measurement outcomes.
\item Converting obtained measurement outcomes into tabulated form.
\end{enumerate}
\subsubsection{Preparing configuration files}
The configuration of PyQBench CLI is driven by YAML files. The first configuration file
describes the experiment scenario to be executed. The second file describes the backend.
Typically, this backend will correspond to the physical device to be benchmarked, but for testing
purposes one might as well use any other Qiskit--compatible backend including simulators.
Let us first describe the experiment configuration file, which might look as follow.
\begin{lstlisting}[language=Python, caption=Defining the experiment, label=lst:experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}
The experiment file contains the following fields:
\begin{itemize}
	\item \texttt{type}: a string describing the type of the experiment. Currently, the only option of  \textit{type} is \texttt{discrimination-fourier}.
	\item \texttt{qubits}: a list enumerating pairs of qubits on which the experiment should be run.
	For configuration in listing \ref{lst:experiment}, the benchmark will run on two pairs of qubits. The first pair is 0 and 1, and
	the second one is 1 and 2. We decided to describe a pair by using \texttt{target} and
	\texttt{ancilla} keys rather than using a plain list to emphasize that the role of qubits in the
	experiment is not symmetric.
	\item \texttt{angles}: an object describing the range of angles for Fourier parameterized
	family. The described range is always uniform, starts at the \texttt{start}, ends at
	\texttt{stop} and contains \texttt{num\_steps} points, including both \texttt{start} and
	\texttt{stop}. The \texttt{start} and \texttt{stop} can be arithmetic expressions using
	\texttt{pi} literal. For instance, the range defined in listing \ref{lst:experiment} contains
	three points: 0, $\pi$ and $2\pi$.
 	 \item \texttt{gateset}: a string describing the set of gates used in the decomposition of
 	 circuits in the experiment. The PyQBench contains explicit implementations of circuits The
 	 possible options are \textit{[ibmq, lucy, rigetti]}, corresponding to decompositions
 	 compatible with IBM Q devices, OQC Lucy device, and Rigetti devices. Alternatively, one might
 	 wish to turn off the decomposition by using a special value \texttt{generic}. However, for this
 	 to work a backend used for the experiment must natively implement all the gates needed for the
 	 experiment, as described in \ref{sec:fourier}.
 	\item \texttt{method}: a string, either \texttt{postselection} or \texttt{direct\_sum}
 	determining which implementation of the conditional measurement is used.
 	\item \texttt{num\_shots}: an integer defines how many shots are performed in the experiment for
 	a particular angle, qubit pair and circuit. Note that if one wishes to compute the total number
 	of shots in the experiment, it is necessary to take into account that the \texttt{postselection}
 	method uses twice as many circuits as the \texttt{direct\_sum} method.
\end{itemize}

The second configuration file describes the backend. We decided to decouple the
experiment and the backend files because it facilitates their reuse. For instance, the same
experiment file can be used to run benchmarks on multiple backends, and the same backend description
file can be used with multiple experiments.

Different Qiskit backends typically require different data for their initialization. Hence, there
are multiple possible formats of the backend configuration files understood by PyQBench. We refer
the interested reader to the PyQBench's documentation. Below we describe an example YAML file
describing IBM Q backend named
Quito.
\begin{lstlisting}[language=Python, caption=IBMQ backend, label=lst:backend]
name: ibmq_quito
asynchronous: false
provider:
	hub: ibm-q
	group: open
	project: main
\end{lstlisting}
IBMQ backends typically require an access token to IBM Quantum Experience. Since it would be unsafe
to store it in plain text, the token has to be configured separately in \texttt{IBMQ\_TOKEN}
environmental variable.

\subsubsection{Remarks on using the asynchronous flag}
For backends supporting asynchronous execution, the \texttt{asynchronous} setting can be configured
to toggle it.
For asynchronous execution to work, the following conditions have to be met:
\begin{itemize}
\item Jobs returned by the backend have unique \texttt{job\_id}.
\item Jobs are retrievable from the backend using the \texttt{backend.retrieve\_job} method, even from another process (e.g. if the original process running the experiment has finished).
\end{itemize}
Since PyQBench cannot determine if the job retrieval works for a given backend, it is the user's
responsibility to ensure that this is the case before setting \texttt{asynchronous} to \texttt{true}.
